---
title: NFL Fantasy Analysis
author: Alexander Cruz
date: '2019-08-19'
slug: nfl-fantasy-analysis
categories:
  - Fantasy Football
tags: []
description: ''
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Packages used:

library(dplyr)

library(readxl)

library(ggplot2)

Data source: https://fantasydata.com/, uses non-ppr standard fantasy scoring
```{r, include=FALSE}
library(dplyr)
library(readxl)
library(ggplot2)
akamara <- "/Users/acruz/My Files/R Project/Evan Engram/Evan.xlsx"
mingram <- "/Users/acruz/My Files/R Project/Evan Engram/Evan.xlsx"

```

Read in Alvin Kamara and Mark Ingram 2018 data, both data sets are in same excel sheets but in different tabs.  Kamara is on sheet 3 and Ingram on sheet 4. Use readxl package, and read_excel() function.
```{r, warning=FALSE, message= FALSE}
Kamara <- read_excel(akamara , sheet = 3)
Mark_Ingram <- read_excel(mingram, sheet = 4)
```
Using dplyr package filter games where A. Kamara had more than 0 rushing attempts and select following columns from dataset: player, week, and points.  Do the same for M. Ingram.
```{r, warning=FALSE, message= FALSE}
Kamara <- Kamara %>% filter(attempt>0) %>% select(player, week, points)
Mark_Ingram <- Mark_Ingram %>% filter(attempt>0) %>% select(player, week)
```
Use dplyr to join the two data sets using function left_join, then create an additional column where it tells you for each week if M. Ingram played, then remove column "player.y".
```{r, warning=FALSE, message= FALSE}
Kamara <- Kamara %>% left_join(Mark_Ingram, by="week") %>% mutate(Ingram_Play= ifelse(is.na(player.y), "Ingram not play", "Ingram played")) %>% select(-player.y)
```
Summarize the data to display difference in fantasy points
```{r, warning=FALSE, message= FALSE}
Kamara1 <- Kamara %>% group_by(player.x,Ingram_Play) %>%
  summarize(`total games`= n(),
            `Average points`= mean(points)) %>% mutate(`Average points` = round(`Average points`,2))
Kamara1

```
Using ggplot2 package display data in bar plot.
```{r, warning=FALSE, message= FALSE}
Kamara2 <- ggplot(data=Kamara1, aes(x=Ingram_Play, y=`Average points`, title="Even")) +
  geom_bar(stat="identity", fill="steelblue") +
  geom_text(aes(label=`Average points`), vjust=10, color="white", size=7) +
  ggtitle("Kamara 2018 Season")
Kamara2

```


