---
title: Get Longitude and Latitude in R
author: Alexander Cruz
date: '2020-05-03'
slug: get-longitude-and-latitude-in-r
categories: []
tags: []
description: ''
---

This tutorial will show you how to retrieve longitude and latitude by simply providing an address.

You will start by installing RDSTK package from github using devtools.  If you don't have devtools, you will also need to install 

```{r, eval=FALSE}
devtools::install_github("rtelmore/RDSTK")
```

Next we will call RDSTK package along with dplyr package

```{r, message=FALSE}
library(RDSTK)
library(dplyr)
```

We will use street2coordinates function to pull lat and long based on location.  The output is a dataframe.  We will use Disneyland location as an example.
```{r} 
location = '1313 Disneyland Dr, Anaheim, CA 92802, United States'
street2coordinates(location)
```
Lets now pull lat and long using dplyr:: select function.  We will use the Los Angeles Staples Center in this example.  
```{r}
df = street2coordinates('1111 S Figueroa St, Los Angeles, CA 90015, United States')
df = df %>% select(full.address, latitude, longitude)
print(df)
```

We can also call multiple addresses in a single function
```{r}
location_example = c('1111 S Figueroa St, Los Angeles, CA 90015, United States', 
                     '1313 Disneyland Dr, Anaheim, CA 92802, United States')
geocode  <- do.call(rbind, lapply(location_example, street2coordinates))
geocode  <- geocode %>% select(full.address, longitude, latitude)
label <- c('Staples Center', 'Disneyland')
geocode <- data.frame(geocode, label=label)
print(geocode)

```

```{r}
library(leaflet)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng= geocode$longitude, lat=geocode$latitude, popup=geocode$label)
m
```

