---
title: Calculate Moving Averages in R
author: Alex Cruz
date: '2022-07-26'
slug: calculate-moving-averages-in-r
categories: []
tags: []
description: ''
---

This tutorial will show you how to calculate moving averages in R.  You will need the zoo and dplyr package to run this script.  
```{r warning= FALSE, message=FALSE}

if(!require(zoo)) install.packages("zoo")
if(!require(dplyr)) install.packages("dplyr")
```

We will create a dataset to use as an example.  The dataset will have date column and a numeric column that we will generate randomly.  We will set a seed so we can recreate the exact numbers later.
```{r warning= FALSE, message=FALSE}
set.seed(7)
start.Date <- as.Date("2022/1/1")
end.Date <- as.Date("2022/4/1")
date.range <- seq.Date(start.Date, end.Date, by = "day")
df <- data.frame("Date" = date.range,
                 "Score" =  runif(length(date.range), 1, 99) )
#show first 10 rows
head(df, n = 10)
```

Now use the rollmean function to calculate your moving average (you can change the range of the moving average by changing the K arugment).  Use the mutate function in order to add rolling average as a new column.
```{r warning= FALSE, message=FALSE}
moving_average <- df %>% mutate(moving_average = rollmean(x = Score, k = 7, fill = NA))
head(moving_average, n = 15)
```

