---
title: Using Match Unique and N_Distinct functions in R
author: Alexander Cruz
date: '2022-09-08'
slug: using-match-unique-and-n-distinct-functions-in-r
categories: []
tags: []
description: ''
---

In this tutorial we will go over using the match, unique, and n_distinct function.  Let's create 3 vectors to use as an example.
```{r}
#Unique, Match, Identical, n_distinct 

setA <- c(5:10)

setB <- c(5,4,7,8,9,11)

setC <- c(5:10)

print(setA)

print(setB)
```

We will use the match function to compare SetA to SetB
```{r}
#using match function
match(setA,setB)
```

how many don't match
```{r}
sum(is.na(match(setA,setB)))
```

Identical will give you a true or false when comparing if both vectors match
```{r}
#using identical 
identical(setA,setB)

identical(setA,setC)
```

n_distinct will count the unqiue values.
```{r}
setD <- c(5,5,6,6,7)
print(setD)

dplyr::n_distinct(setD)

```

Use unique to remove duplicates
```{r}
SetD1 <- unique(setD)
print(SetD1)
```

you can use the names function in combination with identical or match function to check if column names are similar.
```{r}
data("iris")
# head(iris)

iris_1 <- iris
iris_2 <- iris

identical(names(iris_1), names(iris_2))
match(names(iris_1), names(iris_2))
```

Now we will change the first column name to the second dataset and run the same lines of code as above
```{r}
names(iris_2)[1] <- "wrong"
names(iris_2)

identical(names(iris_1), names(iris_2))
match(names(iris_1), names(iris_2))
```

We can remove duplicate rows in a dataframe with the unique function
```{r}
iris4 <- rbind(iris, iris_1)
paste0('numbers of rows: ',nrow(iris4))

iris4 <- unique(iris4)
paste0('numbers of rows after using unique function: ',nrow(iris4))
```



