---
title: How to use Pivot_Wider and Pivot_Longer in R
author: Alex Cruz
date: '2023-04-20'
slug: how-to-use-pivot-wider-and-pivot-longer-in-r
categories: []
tags: []
description: ''
---

First we will call the Tidyverse library and create an example dataset.
```{r, warning=F, message=F, echo=F}
set.seed(7)
library(tidyverse)

x = c(2000, 2000, 2000, 2000, 2000,
      2001, 2001, 2001, 2001, 2001,
      2002, 2002, 2002, 2002, 2002,
      2003, 2003, 2003, 2003, 2003,
      2004, 2004, 2004, 2004, 2004)
df <- data.frame("Color"=c('red', 'blue', 'yellow', 'green', 'orange'),
                 "Score" =  runif(100, 1, 99),
                 "Year" = rep(x, 4))
df1 = df %>% group_by(Year ,Color) %>% summarise(Score = round(sum(Score),0))
```

Let's get to know the dataset.
```{r, warning=F, message=F, echo=T}
head(df1)
names(df1)
n_distinct(df1$Color)
```

Now we will take the Color column and make each color its own colomn.
```{r, warning=F, message=F, echo=T}
wide <- df1 %>% pivot_wider(names_from = Color, values_from = Score)
print(wide)
```

Now we will use pivot longer to change the wide dataset back to its original format
```{r, warning=F, message=F, echo=T}
long <- wide %>% pivot_longer(-Year, names_to = "Color", values_to = "Score" )
print(long)
```

Let's take it a step further and widen out both the year and the color.
```{r, warning=F, message=F, echo=T}
wide2 <- df1 %>% pivot_wider(names_from = c(Year, Color), values_from = Score)
print(wide2)
```

Now we will take the previous dataset (wide2) and revert it back to it's original format
```{r, warning=F, message=F, echo=T}
long2 <- wide2 %>% pivot_longer(everything(), names_sep = "_", names_to = c("Year","Color"), values_to = "Score" )
print(long2)
```
